---
title: "DMLMbvs-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DMLMbvs-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(DMLMbvs)
```
This document contains a tutorial for the R package DMLMbvs, as described in *A Bayesian model of microbiome data for simultaneous identification of covariate associations and prediction of phenotypic outcomes,* by M.D. Koslovsky, K. Hoffman, C. Daniel-MacDougall, and M. Vannucci. The code below can be used to replicate the simulation study and sensitivity analysis presented in the main manuscript. 

# Simulate Data
Having installed and loaded the `DMLMbvs` package into the R environment, generate a simulated data set using the `simulate_DMLM()` function. 

```{r }
data <- simulate_DMLM()
```

By default 50 subjects (`subject_sim = 50`) with 50 taxa (`B_sim = 50`) and 50 covariates (`covariates_sim = 50`) are simulated. Of these, 10 covariates (`active_cov = 10`) are associated with the taxa counts, and the first 10 balances are associated with the continous outcome. The covariates are simulated from a normal distribution with covariance set to 0.4 as default (`covar = 0.4`). Balances are constructed using sequential binary separation. 

# Run model
To run the model, simply imput the data to the `dm_lm_bvs_R()` function.  

```{r , message = FALSE, results = 'hide' }
output <- dm_lm_bvs_R( y = data$Y, z = data$Z, x = data$X )
```
By default, the algorithm is run for 10,000 iterations, thinning to every 10$^{th}$ iteration. The prior probabilities of inclusion for covariates and balances are both 10%. See the R documentation for various options involving the default parameterization. The `output` object contains a list (in order) of the MCMC samples for $\alpha$, $\zeta$, $\phi$, $\psi$, and $\xi$, following the notation used in the main manuscript. Using the `selected()` function, the user can extract the selected covariates and balances in the model. By default, the threshold for active terms is 0.5. For example, 
```{r }
select <- selected( output, burnin = 500 )
```

The `select` object contains the selected covariates and balances, as well as the marginal posterior probabilities of inclusion (MPPI) for all of the terms. We can see that the model does a good job recovering the active terms in this example:
```{r }
# Selected covariate-taxon relations
select$selected_zeta

# True covariate-taxon relations
data$true_cov

# Selected balances (First 10 active in true model)
select$selected_xi
```

For inference, the `selected()` function can be used to generate plots of the MPPIs for both levels of the model as well as plots for the number of terms active in the model at each MCMC iteration. The latter can be used to help assess convergence of the model. To obtain these plots, simply run 

```{r}
select <- selected( output, burnin = 500, plotting = TRUE )
```

To control for multiplicity, a Bayesian false discovery rate threshold can be used. To determine this threshold (and the corresponding selected terms), use the `bfdr_wrap()` function and input the vector of MPPIs obtained with the `selected()` function. 
```{r }
bfdr_wrap( select$mppi_zeta)

bfdr_wrap( select$mppi_xi)
```

This function outputs the threshold and number of covariates selected for a set of Bayesian false discovery rates. For example to determine the selected terms with a 5% BFDR on each level of the model seperately, run `selected( output, c( 0.834, 0.934 ), burnin = 500 )`.

